---
title: "Assignment 02 - Data Viz and Wrangling"
author: "Robin Du"
output: github_document
always_allow_html: true
---

# Due Date

September 23, 2020 by midnight Pacific time.

For this assignment, we will be analyzing data from USC’s Children’s Health Study. The learning objectives are to conduct data wrangling and visualize the data with key questions in mind.

# Data Wrangling

You will need to download two datasets from https://github.com/USCbiostats/data-science-data. The individual and regional CHS datasets in 01_chs. The individual data includes personal and health characteristics of children in 12 communities across Southern California. The regional data include air quality measurements at the community level. Once downloaded, you can merge these datasets using the location variable.

```{r setup}
library(data.table)
library(dplyr)
individual <- fread("C:/PM566/HW2/chs_individual.csv")
regional <- fread("C:/PM566/HW2/chs_regional.csv")

# Check dimensions of datasets
dim(individual)
dim(regional)

# Check variable names
individual[1:5,]
regional[1:5,]
```

- There are 1200 rows and 23 columns in the individual dataset
- There are 12 rows and 27 columns in the regional dataset

- Both datasets have a variable called "townname"


```{r merging-datasets}
chs <- merge(
  x = individual, y = regional,
  by.x = "townname", by.y = "townname",
  all.x = TRUE, all.y = FALSE
)

dim(chs)
chs[1:5,]
```

- There are 1200 rows (same as the individual dataset) and 49 columns in the merged dataset


```{r imputing}
# Impute missing data using the average within the variables "male" and "hispanic"
chs[, bmi_mean := mean(bmi, na.rm = TRUE), by = .(male, hispanic)]
chs[, bmi := ifelse(is.na(bmi), bmi_mean, bmi)]

chs[, smoke_mean := mean(smoke, na.rm = TRUE), by = .(male, hispanic)]
chs[, smoke := ifelse(is.na(smoke), smoke_mean, smoke)]

chs[, gasstove_mean := mean(gasstove, na.rm = TRUE), by = .(male, hispanic)]
chs[, gasstove := ifelse(is.na(gasstove), gasstove_mean, gasstove)]

chs[, fev_mean := mean(fev, na.rm = TRUE), by = .(male, hispanic)]
chs[, fev := ifelse(is.na(fev), fev_mean, fev)]

chs[, asthma_mean := mean(asthma, na.rm = TRUE), by = .(male, hispanic)]
chs[, asthma := ifelse(is.na(asthma), asthma_mean, asthma)]
```


Once combined, you will need to do the following.

In the case of missing values, impute data using the average within the variables “male” and “hispanic.” If you are interested (and feel adventurous) in the theme of Data Imputation, take a look at this paper on “Multiple Imputation” using the Amelia R package here.


## 1: After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r question1}
nrow(chs)
nrow(individual)
```
- Again, the number of rows in the merged chs dataset is the same as the number of rows in the individual dataset; there are no duplicates


## 2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r question2}
chs[, obesity_level := cut(chs$bmi,
                           breaks=c(-Inf,14,22,24,Inf),
                           labels=c("underweight","normal","overweight","obese"))]

chs %>% group_by(obesity_level) %>%
summarise(
          min_bmi = min(bmi, na.rm = TRUE),
          max_bmi = max(bmi, na.rm = TRUE),
          count = n()
          ) %>%
filter(!is.na(obesity_level))
```


## 3. Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r question3}
chs$smoke_gas_exposure[chs$smoke==0 & chs$gasstove==0] <- "no exposure"
chs$smoke_gas_exposure[chs$smoke==0 & chs$gasstove==1] <- "gas exposure only"
chs$smoke_gas_exposure[chs$smoke==1 & chs$gasstove==0] <- "smoke exposure only"
chs$smoke_gas_exposure[chs$smoke==1 & chs$gasstove==1] <- "both smoke and gas exposure"

head(chs)
```



## 4. Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r question4}
# By Town
chs %>% group_by(townname) %>%
  summarise(
    fev_avg = mean(fev, na.rm = TRUE),
    fev_sd = sd(fev, na.rm = TRUE),
    asthma_prop = sum(asthma, na.rm = TRUE)/length(asthma)
  )

# By Sex
chs %>% group_by(male) %>%
  summarise(
    fev_avg = mean(fev, na.rm = TRUE),
    fev_sd = sd(fev, na.rm = TRUE),
    asthma_prop = sum(asthma, na.rm = TRUE)/length(asthma)
  )

# By Obesity Level
chs %>% group_by(obesity_level) %>%
  summarise(
    fev_avg = mean(fev, na.rm = TRUE),
    fev_sd = sd(fev, na.rm = TRUE),
    asthma_prop = sum(asthma, na.rm = TRUE)/length(asthma)
  )

# By Smoke_gas_exposure
chs %>% group_by(smoke_gas_exposure) %>%
  summarise(
    fev_avg = mean(fev, na.rm = TRUE),
    fev_sd = sd(fev, na.rm = TRUE),
    asthma_prop = sum(asthma, na.rm = TRUE)/length(asthma)
  )
```


# Looking at the Data (EDA)

The primary questions of interest are:
1. What is the association between BMI and FEV (forced expiratory volume)?
2. What is the association between smoke and gas exposure and FEV?
3. What is the association between PM2.5 exposure and FEV?

Follow the EDA checklist from week 3 and the previous assignment. Be sure to focus on the key variables. Create the following figures and interpret them. Be sure to include easily understandable axes, titles, and legends.

```{r eda-setup}
# Check the dataset
dim(chs)
head(chs)
tail(chs)
str(chs)

# Examine the main variables of interest
summary(chs$bmi)
summary(chs$fev)
summary(chs$pm25_mass)
table(chs$smoke_gas_exposure)
```
- The missing data were accounted for by imputing (see Data Wrangling section). The summary of the main variables seems reasonable.


## 1. Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.

```{r}
library(ggplot2)
chs %>%
  filter(!(townname %in% NA)) %>%
ggplot(mapping = aes(x = fev, y = bmi)) +
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x) + 
  facet_wrap(~townname) +
  labs(title = "Regression of BMI vs FEV by Town")
```

- It appears that there is a positive association between BMI and FEV for every town


## 2. Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.

```{r}
# By BMI category
chs %>%
  filter(!(obesity_level %in% NA)) %>%
ggplot(mapping = aes(x = fev, fill = obesity_level)) +
  geom_histogram(binwidth = 30) +
  labs(title = "FEv by BMI Category") +
  scale_fill_brewer(palette = "Set1")

# By Smoke/gas exposure category
chs %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(mapping = aes(x = fev, fill = smoke_gas_exposure)) +
  geom_histogram(binwidth = 30) +
  labs(title = "FEv by Smoke/Gas Exposure") +
  scale_fill_brewer(palette = "Set1")
```

- The histograms show that people in the higher BMI categories tend to have higher FEV, while smoke/gas exposure does not seem to have an association with FEV


## 3. Barchart of BMI by smoke/gas exposure.

```{r}
chs %>%
  filter(!(obesity_level %in% NA)) %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(mapping = aes(x = obesity_level, fill = smoke_gas_exposure)) +
  geom_bar() +
  labs(title = "BMI by Smoke/Gas Exposure") +
  scale_fill_brewer(palette = "Set1")
```

- There appears to be a fairly proportional spread of smoke/gas exposure among the BMI categories, indicating that there is little/no association between smoke/gas exposure and BMI category


## 4. Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.

```{r}
# By BMI category
chs %>%
  filter(!(obesity_level %in% NA)) %>%
ggplot(mapping = aes(x = obesity_level, y = fev)) +
  stat_summary(
       fun.min = min,
       fun.max = max,
       fun= median) +
    labs(title = "Statistical Summary Graphs of FEV by BMI Category")

# By Smoke/gas exposure category
chs %>%
  filter(!(smoke_gas_exposure %in% NA)) %>%
ggplot(mapping = aes(x = smoke_gas_exposure, y = fev)) +
  stat_summary(
       fun.min = min,
       fun.max = max,
       fun= median) +
    labs(title = "Statistical Summary Graphs of FEV by Smoke/Gas Exposure Category")
```


- It appears that FEV increased with BMI category, but not with smoke/gas exposure category (or very little)


## 5. A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.

```{r}

```


## 6. Choose a visualization to examine whether PM2.5 mass is associated with FEV.

```{r}

```

