---
title: "Assignment 02 - Data Viz and Wrangling"
author: "Robin Du"
output: github_document
always_allow_html: true
---

# Due Date

September 23, 2020 by midnight Pacific time.

For this assignment, we will be analyzing data from USC’s Children’s Health Study. The learning objectives are to conduct data wrangling and visualize the data with key questions in mind.

# Data Wrangling

You will need to download two datasets from https://github.com/USCbiostats/data-science-data. The individual and regional CHS datasets in 01_chs. The individual data includes personal and health characteristics of children in 12 communities across Southern California. The regional data include air quality measurements at the community level. Once downloaded, you can merge these datasets using the location variable.

```{r setup}
library(data.table)
individual <- fread("C:/PM566/HW2/chs_individual.csv")
regional <- fread("C:/PM566/HW2/chs_regional.csv")

# Check dimensions of datasets
dim(individual)
dim(regional)

# Check variable names
individual[1:5,]
regional[1:5,]
```

- There are 1200 rows and 23 columns in the individual dataset
- There are 12 rows and 27 columns in the regional dataset

- Both datasets have a variable called "townname"


```{r merging-datasets}
chs <- merge(
  x = individual, y = regional,
  by.x = "townname", by.y = "townname",
  all.x = TRUE, all.y = FALSE
)

dim(chs)
chs[1:5,]
```

- There are 1200 rows (same as the individual dataset) and 49 columns in the merged dataset


Once combined, you will need to do the following:


## 1: After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.

```{r}
nrow(chs)
nrow(individual)
```
- Again, the number of rows in the merged chs dataset is the same as the number of rows in the individual dataset; there are no duplicates


## 2. Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r}

```

